---
title: "summary_of_session_2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As always, we started by loading in the tidyverse.

```{r, message=FALSE}
library(tidyverse)
```

# Reading in the data

The cps data we aim to analyze is a dta file. We weren't quite sure how to read that in so we googled read in stata files into r and then added "hadley" to the search and filtered to only see results in the last 1 year.

We found our solution in the `haven` package:

```{r}
library(haven)

cps <- read_dta(here::here("data-raw", "cps_mar_2000_new with additional vars.dta"))
```

We also noticed that there is a way to look at the variable labels:

```{r}
print_labels(cps$sex)
```

# Cleaning the data

We played around with cleaning the data and eventually arrived at the following:

```{r}
cps_new <-
    cps %>%
    sample_frac(0.01) %>%
    select(incwage, age, sex, yrsed) %>%
    mutate(
        sex = case_when(sex == 1 ~ "male", sex == 2 ~ "female"),
        college = case_when(yrsed >= 13 ~ "college", yrsed <= 12 ~ "no_college")
    ) %>%
    filter(age >= 16)
```

# Plotting the data

We walked through building the following plot:

```{r}
cps_new %>%
    ggplot(aes(x = age, y = incwage)) +
    geom_point(alpha = 0.25) +
    facet_grid(sex ~ college) +
    geom_smooth(method = "lm")
```

# Modeling the data

Lastly, we built a linear model 

```{r}
mod <- lm(incwage ~ age + college + sex, data = cps_new)
```

Then we looked at the model using the old way with `summary`.

```{r}
summary(mod)
```

And a newer way using the `broom` package.

```{r}
library(broom)

tidy(mod)
```

We noticed that the tidy function returns a data frame which we know how to work with:

```{r}
tidy(mod) %>%
    filter(p.value <= 0.05)
```

Lastly, we used the `modelr` package to show how to use the model to add predictions and residuals to our data:

```{r, message=FALSE}
library(modelr)

cps_new %>%
    add_predictions(mod) %>%
    add_residuals(mod)
```
